---
title: Parse the primary, district judges elections at county level
author: Duncan Garmonsway
output: html_document
---

First, run `/R/download.R` to download the source files.

```{r, echo = TRUE}
library(tidyverse)
library(stringr)
library(stringi)
library(tidyxl)
library(unpivotr)
library(here)

sources_path <- here("sources")
output_path <- here("output")
xlsx_dir <- here("xlsx")
csv_dir <- here("csv")
idaho_home <- "http://www.sos.idaho.gov/ELECT/results/"
idaho_url <- paste0(idaho_home, "index.html")
url_table_path <- file.path(sources_path, "url-table.html")
working_dir <- here("working")
files_path <- file.path(working_dir, "files.csv")
county_files_path <- file.path(working_dir, "county_files.csv")
```

The various kinds of files to work with.  This script handles statewide,
primary, county.  The numbers are the number of files of the given kind.

```{r, echo = TRUE}
files <- read_csv(files_path)
files %>%
  select(scope, election, geography) %>%
  ftable(row.vars = 1:2, col.vars = 3)
```

A function to import the data from a single table on a single sheet, where each
district is treated as a different table.

```{r, echo = TRUE}
district_votes <- function(district_chunk, fillgrey, bold) {
  district <-
    district_chunk %>%
    filter(col == 1, row == min(row)) %>%
    pull(character)
  cat("district: ", district, "\n")
  candidates <-
    district_chunk %>%
    filter(col == 1, !is.na(character), !(local_format_id %in% bold)) %>%
    mutate(candidate = character) %>%
    select(row, col, candidate)
  counties <-
    district_chunk %>%
    filter(row == min(row), col != 1, !is.na(character)) %>%
    mutate(county = character) %>%
    select(row, col, county)
  district_chunk %>%
    filter(!is.na(numeric)) %>%
    mutate(votes = numeric) %>%
    select(row, col, votes) %>%
    semi_join(candidates, by = "row") %>%
    semi_join(counties, by = "col") %>%
    W(candidates) %>%
    N(counties) %>%
    mutate(district = district) %>%
    select(county, district, candidate, votes)
}
```

A function to import the data for all districts on a single sheet.

```{r, echo = TRUE}
sheet_votes <- function(name, cells, fillgrey, bold) {
  cat("sheet: ", name, "\n")
  districts <-
    cells %>%
    arrange(row, col) %>%
    mutate(chunk = cumsum(col == 1 & local_format_id %in% fillgrey)) %>%
    split(.$chunk) %>%
    .[-1]
  map_df(districts, district_votes, fillgrey, bold)
}
```

A function to import the data for all sheets in a file.

```{r, echo = TRUE}
book_votes <- function(path) {
  cat("book: ", path, "\n")
  book <- tidy_xlsx(path)
  sheets <- book$data
  formats <- book$formats
  fillgrey <- which(!is.na(formats$local$fill$patternFill$bgColor$rgb))
  bold <- which(formats$local$font$bold)
  map2_df(names(sheets), sheets, sheet_votes, fillgrey, bold)
}
```

Apply the above functions to every legislature , primary, county file.

```{r, echo = TRUE}
votes <-
  files %>%
  filter(scope == "district judges",
         election == "primary",
         geography == "county") %>%
  pull(xlsx_path) %>%
  set_names(.) %>%
  map_df(book_votes, .id = "book") %>%
  select(county, district, candidate, votes)
```

Check the data quality by listing unique values in each field. Not perfect, and
beyond my expertise to map these values to standard, correct ones.

```{r, echo = TRUE}
votes %>%
  select(-votes) %>%
  map(~ .x %>%
      str_replace("\\n", " ") %>%
      str_trim() %>%
      table())
```

Write to output file.

```{r, echo = TRUE}
votes %>%
  write_csv(file.path(output_path, "yyyymmdd__id__primary__county__judges.csv"))
```
