---
title: Parse the primary, statewide elections at county level
author: Duncan Garmonsway
output: html_document
---

First, run /R/download.R to download the source files.

```{r, echo = TRUE}
library(tidyverse)
library(stringr)
library(stringi)
library(tidyxl)
library(unpivotr)
library(here)

sources_path <- here("sources")
output_path <- here("output")
xlsx_dir <- here("xlsx")
csv_dir <- here("csv")
idaho_home <- "http://www.sos.idaho.gov/ELECT/results/"
idaho_url <- paste0(idaho_home, "index.html")
url_table_path <- file.path(sources_path, "url-table.html")
working_dir <- here("working")
files_path <- file.path(working_dir, "files.csv")
county_files_path <- file.path(working_dir, "county_files.csv")
```

The various kinds of files to work with.  This script handles statewide,
primary, county.  The numbers are the number of files of the given kind.

```{r, echo = TRUE}
files <- read_csv(files_path)
files %>%
  select(scope, election, geography) %>%
  ftable(row.vars = 1:2, col.vars = 3)
```

Functions to pull the district number from strings like:

* "REPRESENTATIVE DISTRICT 1"
* "Representative District 2"
* "1st DISTRICT REPRESENTATIVE"
* "2nd District Representative"

```{r, echo = TRUE}
get_district <- function(x) {
  map_dbl(x,
          function(y) {
            if (str_detect(y, "[0-9]")) {
              y %>%
                str_replace_all("\\.", "") %>%
                parse_number()
            } else {
              NA
            }
          })
}
forget_district <- function(x) {
  str_replace_all(x, " [0-9]+.*$", "")
}
```

A function to import the data from a single table on a single sheet (98prcnty.xlsx
has two tables on the same sheet)

```{r, echo = TRUE}
table_votes <- function(corner, cells, bold, top_border) {
  county <-
    corner %>%
    offset_S(cells, 1) %>%
    extend_S(cells, boundary = ~ str_detect(character, "TOTAL")) %>%
    filter(!is.na(character)) %>%
    mutate(county = str_trim(character)) %>%
    select(row, col, county)

  party <-
    cells %>%
    filter(row == corner$row - 1,
           col != 1,
           local_format_id %in% unlist(top_border),
           !is.na(character)) %>%
    mutate(party = str_trim(character)) %>%
    select(row, col, party)

  if (nrow(party) == 0) party <- data_frame(row = 1, col = 1, party = NA)

  candidate <-
    cells %>%
    filter(row == corner$row, col != 1,
           !is.na(character)) %>%
    mutate(candidate = str_trim(character)) %>%
    select(row, col, candidate)

  office <-
    cells %>%
    filter(row < corner$row,
           row > corner$row - 5,
           col != 1,
           local_format_id %in% unlist(bold),
           !is.na(character)) %>%
    mutate(office = str_trim(character)) %>%
    select(row, col, office) %>%
    group_by(col) %>%
    arrange(col, row) %>%
    summarise(office = paste(office, collapse = " ")) %>%
    mutate(district = get_district(office),
           office = forget_district(office)) %>%
    mutate(row = 1L) %>%
    select(row, col, office, district)

  statistics_col <-
    office %>%
    filter(str_detect(office, "STATISTIC")) %>%
    pull(col) %>%
    min(Inf)

  votes <-
    cells %>%
    semi_join(county, by = "row") %>%
    semi_join(candidate, by = "col") %>%
    filter(col < statistics_col,
           !is.na(content)) %>%
    mutate(votes = parse_number(content)) %>%
    select(row, col, votes)

  votes %>%
    W(county) %>%
    NNW(office) %>%
    NNW(party) %>%
    N(candidate) %>%
    select(county, office, district, party, candidate, votes)
}
```

A function to iterate over many tables on a single sheet (98prcnty.xlsx has two
tables on the same sheet)

```{r, echo = TRUE}
sheet_votes <- function(cells, name, bold, top_border) {
  if (name %in% c("sheet1")) {
    # There's one of these in 04_pri_cnty.xlsx, and it can be skipped because it
    # is simply 'working' for the actual data in the previous sheet.
    cat("  skipping sheet: ", name, "\n")
    return(data_frame(county = character(),
                      office = character(),
                      district = character(),
                      party = character(),
                      candidate = character(),
                      votes = numeric()))
  }
  cat("  sheet: ", name, "\n")

  # The cell with the string 'counties' is reliably the top-left corner cell of
  # a table.
  counties <-
    cells %>%
    filter(col == 1,
           str_trim(tolower(character)) == "counties") %>%
    split(.$address)

  map_df(counties, table_votes, cells, bold, top_border)
}
```

A function to iterate over all the sheets in a workbook.

```{r, echo = TRUE}
book_votes <- function(x) {
  cat("book: ", x, "\n")
  x <- tidy_xlsx(x)
  top_border <- list(which(!is.na(x$formats$local$border$top$style)))
  bold <- list(which(x$formats$local$font$bold))
  pmap_df(list(x$data,
               str_trim(tolower(names(x$data))),
               bold,
               top_border),
          sheet_votes, .id = "sheet")
}
```

Apply the above functions to every statewide, primary, county file.

```{r, echo = TRUE}
votes <-
  files %>%
  filter(scope == "statewide",
         election == "primary",
         geography == "county") %>%
  pull(xlsx_path) %>%
  set_names(.) %>%
  map_df(book_votes, .id = "book")
```

Check the data quality by listing unique values in each field. Not perfect, and
beyond my expertise to map these values to standard, correct ones.

```{r, echo = TRUE}
votes %>%
  select(-votes) %>%
  map(~ .x %>%
      str_replace("\\n", " ") %>%
      str_trim() %>%
      unique() %>%
      sort())
```

Write to output file.

```{r, echo = TRUE}
votes %>%
  write_csv(file.path(output_path,
                      "yyyymmdd__id__primary__county__statewide.csv"))
```
